---
title: "Glover Final Model"
output:
  pdf_document: default
  html_document: default
date: "2024-10-18"
---

```{r}
tinytex::install_tinytex(force = TRUE)

library(tidyverse)
library(dplyr)
library(here)
library(rethinking)
library(cmdstanr)

library(readxl)
SDAmaster <- read_excel("C:/Users/19glo/OneDrive/Desktop/Bayesian/SDAmaster.xlsx")

```


# My data has to do with the increase in metabolic rate (measured in my case by change in CO2 production) over time since digestion. This week I made the hierarchical model, tried to compare it to my old model and then extracted a bunch of interesting datapoints from this model. 

```{r}

SDAmaster <- SDAmaster %>%
  mutate(Species_TrialType = paste(Species, TrialType, sep = "_"))


ggplot(aes(x=Time_Actual, y=Average_VCO2,  color = Species_TrialType), data= SDAmaster) +
  geom_smooth() +
  labs(title="VCO2 Over Time 7p", x="Time", y="VCO2")

```




#Remember as you're seeing this that I don't have a full dataset yet, so there is still a lot of data that I still need to fill out xp and control. In this graph, 7p represents a meal which is 7 give or take 0.5 percent chameleon body mass, xp is an meal which is lower than 7 percent, and control is a fasted chameleon. 


#In this section I standardize my variables and create a dataset which only includes the variables that I want.

```{r}
#Here I've made a new index variable which includes both Species and TrialType
sd_tidy = SDAmaster %>%
  mutate(SpeciesTrialType = paste(Species, TrialType, sep = "_"))


#This makes a dataset with all of the variables + New standardized Variables
sd_tidy  = sd_tidy %>% 
  mutate(T2 = standardize(Time_Actual),
         STT = as.integer(as.factor(SpeciesTrialType)),
         CO2 = standardize(Average_VCO2),
         Species1 = as.integer(as.factor(Species)),
         TrialType1 = as.integer(as.factor(TrialType)),
         CPercent = standardize(Cricket_Percent),
         TrialType,
         Species)

#This makes a much smaller dataset with only what you need which keeps the model from running the entire data set and lowers the amount of time it takes

sd_short = sd_tidy %>% dplyr::select(CO2, Species1, Species, T2, TrialType1, CPercent, TrialType, STT)
```



#This is my old model which includes the STT index variable to predict changes in CO2 over time since feeding given species and trial type. It includes a T^2 variable because I expect the shape of the curve to be an upsidedown parabola.

```{r}
CO2_Time3 <-ulam(
alist(
  CO2 ~ dnorm(mu,sigma),
  #This includes an interaction between Time and Time squared since it is likely an upside down parabola. Each time you add the bracketed things it adds another line for that category
  mu <- a + b1[STT] + b2[STT] * T2 + b3[STT] * T2^2,          
  a ~ dnorm(0, 1),
  b1[STT] ~ dnorm(0, 1),
  b2[STT] ~ dnorm(0, 1),
  b3[STT] ~ dnorm(0, 1),
 # b4 ~ dnorm(0, 1),
  sigma ~ dexp(1)
) ,data = sd_short)
```


#This is my new model which is hierarchical. I kept the same basic equation and parameters with the addition that each variable parameter is predicted by the multinormal function including a, b, Rho, and sigma_id. Since the data is standardized, I just used a basic normal distribtions of (0,1) for a and b, and then exponential for sigma_id. Rho needs to deal with the matrix as seen above, so it is necessary to use the lkj_corr function. 

```{r}
CO2_Time4 <-ulam(
alist(
  CO2 ~ dnorm(mu,sigma),
  #This includes an interaction between Time and Time squared since it is likely an upside down parabola. Each time you add the bracketed things it adds another line for that category
  mu <- a1[STT] + b1[STT] * T2 + b2[STT] * T2^2 ,
  c(a1, b1, b2)[STT] ~ multi_normal(c(a, b), Rho, sigma_id),
  a ~ dnorm(0, 1),
  b ~ dnorm(0, 1),
  sigma_id ~ dexp(1),
  Rho ~ lkj_corr(2),
  sigma ~ dexp(1)
) ,data = sd_short, chains = 4, cores = 4)
```



# Since I can't seem to get the compare() function to work I just put both of the precis outputs here. Notable differences include a large positive value in model 3 for STT[1] compared to the large and distant from 0 negative value in model 4. I'm not sure why model 3 did that because it makes much more sense that it would be negative for the shape. Ess_bulk is increased by an order of magnitude as well in the model 4 which indicates that there is much better mixing and that the model is taking more from the datapoints and iterations than did model 3. rhat is also notably right at 1.00 which is better than model 3 which ranges a little bit. 

# For ease of understanding: STT 1 = Brygooi 7p, 2 = Brygooi Control after 7p 3 = Brygooi Control after xp, 4 = Brygooi xp, 5 = Thieli 7p, 6 = Thieli Control after 7p, 7 = Thieli Control after xp, 8 = Thieli xp. I have the most data for Brygooi 7p and Thieli xp. I went ahead and broke apart the controls by the size of their last meal since I think there might be a difference. 

# Just to describe the outputs of the precis for model 4 (since it's the new model), I notice that Brygooi 7p has a large (far from 0) negative value for the time squared coefficient , a positive value for the linear coefficient and a much higher intercept a[1] value which is quite far from zero. All of this makes me think that this is a decent model, and we can feel pretty sure that it differs from the other indexes. I'll go into the CI stuff on later extracted values but not here since it's pretty difficult to actually interpret this given it's a quadratic function. Also of note is STT[2] and STT[3] which seem to have their coefficients near zero with relatively large standard deviations. I would interpret this as meaning that the lines are pretty flat, especially as compared to STT[1]. Interestingly, they differ from eachother in that a1[2] is negative and far from zero where a1[3] is right around zero. They also differ from Thieli where the ones with the most data a1[2] vs a1[7] do not overlap and the Thieli control is much lower than the Brygooi (-.87 vs -.33 means). This makes me feel confident that there is a clear difference between the two. In the future I'll have to include a variable which accounts for chameleon body mass which may be confounding the difference in species here. STT[8] which is Thieli xp looks to be slightly negative and does not quite include 0 which makes sense for its predicted shape. There is not much in terms of effect from the linear coefficient, but like the control for this species, also has a relatively low intercept relative to brygooi. The last thing I'll note here is that the control for thieli does not have a large effect from the quadratic coefficient but has a large negative linear effect which is interesting. I wonder if they are maybe still metabolizing the food even several days after feeding. 

```{r}
precis(CO2_Time3, depth = 2)
precis(CO2_Time4, depth = 3)
```


# Here I plot the data by category 1 which is Brygooi 7p (the one I have the most data for). You can see that there is a of parabolic shape going on, and it looks about how I'd expect. I wanted to check to see how it compares against the new model. I am a bit confused though since the precis predicts a positive coefficient for the quadratic part, but this line is clearly an upsidedown parabola.   

```{r}
#This creates the x axis of the plot
t2_seq = seq(-1.3, 2.6, length.out = 1000)

#this samples a bunch of predictions for the coefficients of the model given the data and the parameters. This is only for STT = 1 or Brygooi 7p.
mu_pred <- link(CO2_Time3, data = data.frame(T2 = t2_seq,
                                       STT = 1))

#This takes our predictions and creates a line based on the mean of the link output
mu_mean <- apply(mu_pred, 2, mean)

#This takes our samples predictions and produces our credible interval at 95%
mu_PI <- apply(mu_pred, 2, PI, prob = 0.95)

#This plots only the values for the value of trialtype that you want. In this case 7p is 1, control is 2, and xp is 3.

plot(x = sd_short %>% filter(STT == 1) %>% pull(T2), 
     y = sd_short %>% filter(STT == 1) %>% pull(CO2), 
     col = col.alpha(rangi2, 0.5), 
     xlab = "Time", 
     ylab = "VCO2", 
     pch = 16)

#These create the line and confidence interval on the plot
lines(t2_seq, mu_mean, col = "blue", lwd = 2)
shade_color <- col.alpha("black", 0.3)
shade(mu_PI, t2_seq, col=shade_color)
```

# Here is the newer model and plot. Not sure what I was expecting, but these look almost exactly the same. They both do seem to follow the general trend well though, so I'm happy about that. 

```{r}
#This creates the x axis of the plot
t4_seq = seq(-1.3, 2.6, length.out = 1000)

#this samples a bunch of predictions for the coefficients of the model given the data and the parameters. This is only for STT = 1 or Brygooi 7p.
mu_pred4 <- link(CO2_Time4, data = data.frame(T2 = t4_seq,
                                       STT = 1))

#This takes our predictions and creates a line based on the mean of the link output
mu_mean4 <- apply(mu_pred4, 2, mean)

#This takes our samples predictions and produces our credible interval at 95%
mu_PI4 <- apply(mu_pred4, 2, PI, prob = 0.95)

#This plots only the values for the value of trialtype that you want. In this case 7p is 1, control is 2, and xp is 3.

plot(x = sd_short %>% filter(STT == 1) %>% pull(T2), 
     y = sd_short %>% filter(STT == 1) %>% pull(CO2), 
     col = col.alpha(rangi2, 0.5), 
     xlab = "Time", 
     ylab = "VCO2", 
     pch = 16)

#These create the line and confidence interval on the plot
lines(t4_seq, mu_mean4, col = "blue", lwd = 2)
shade_color <- col.alpha("black", 0.3)
shade(mu_PI4, t4_seq, col=shade_color)
```


# In the next few parts I extract some important datapoints for my model. This is Brygooi 7p which has the most stark peak and should show up the best. After some data wrangling help from Jeff, we were able to take a bunch of iterations with very small spaces (t4_seq has length.out = 1000) to predict a peak value and the time at which it occurs. Given the outputs here, we can say that we are 95% confident that the peak VCO2 production for Brygooi 7p is between 4.559 and 10.578 mL per hour. We are also 95% confident that the peak occurs between minutes 986 and 1047. I'll comparge these values to others below.

# A quick note here these values may change slightly between now and when I turn it in since I keep adding data to the dataset, so if I miss some or forget to change them before I turn it in that is why the numbers I use to extract the points might differ slightly from the outputs of the tibbles. 

```{r}

#This creates a nice tibble where it takes all of the mu_preds from a link function on the model, arranges the data, and then filters out where each iteration hits its peak. It then takes the mean time value where it hits that peak and the standard deviation of those peak times

library(tidyverse)
times = tibble(t4_seq = t4_seq) %>% 
  mutate(name = paste0("V", row_number()))
mu_pred4 %>% as_data_frame() %>% 
  mutate(iter = 1:nrow(.)) %>% 
  pivot_longer(cols = -iter) %>% 
  left_join(times) %>% 
  group_by(iter) %>% 
  filter(value == max(value)) %>% 
  ungroup %>% 
  reframe(mean = mean(t4_seq),
          sd = sd(t4_seq))


#this samples a bunch of predictions for the coefficients of the model given the data and the parameters. This is only for STT = 1 or Brygooi 7p.
mu_pred4 <- link(CO2_Time4, data = data.frame(T2 = t4_seq,
                                       STT = 1))

#This takes our predictions and creates a line based on the mean of the link output
mu_mean4 <- apply(mu_pred4, 2, mean)

#This takes our samples predictions and produces our credible interval at 95%
mu_PI4 <- apply(mu_pred4, 2, PI, prob = 0.95)


# THis tibble takes all of the predicted lines and averages them at each of the 1000 timepoints from t4_seq. It then arranges them so that the top value is the largest or the peak and gives the standard deviation.
tibble(mean = mu_mean4,
       lower = mu_PI4[1,],
       upper = mu_PI4[2,],
       sd = sd(mu_mean4),
       time = t4_seq) %>% 
  arrange(-mean)


# In these lines I take the values for the mean peak VCO2 and its standard deviation and destandardize them. I then used the 95% CI which is 1.96 stdevs from the mean. 
mean_VO21 <- 1.700777 * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)
upper_VO21 <- (1.700777 + 1.96 * (.7674976)) * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)
lower_VO21 <- (1.700777 - 1.96 * (.7674976)) * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)

mean_VO21
upper_VO21
lower_VO21


# In these lines I take the mean and sd output from the first tibble that I created and destandardize those values. These correspond to the timepoint and 95% CI that the curve reaches its peak. 
mean_time1a <- .490607 * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)
upper_time1a <- (.490607 + 1.96 * (.027397)) * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)
lower_time1a <- (.490607 - 1.96 * (.027397)) * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)


mean_time1a
upper_time1a
lower_time1a
```




#Here I do the same but for the Brygooi xp which doesn't have as much data as the other but should produce some similar shape. The predicted range for the peak value in this case is lower on average where we are 95% confident that it is between 3.57 and 6.11. Given this, we can feel confident that there is a difference between the means for 7p and xp. The timepoints also differ in that it took longer to reach the peak in this case where our 95% CI is between 1081 and 1734. While this is an admittedly large range, it doesn't even inlcude the values from the 7p curve which is interesting. I haven't read anything about the time to peak changing based on the amount of food eaten in other species. I'll have to collect more data to determine if this is a real pattern. 

```{r}
mu_pred3a <- link(CO2_Time4, data = data.frame(T2 = t4_seq,
                                       STT = 4))
mu_mean3a <- apply(mu_pred3a, 2, mean)
mu_PI3a <- apply(mu_pred3a, 2, PI, prob = 0.95)

times = tibble(t4_seq = t4_seq) %>% 
  mutate(name = paste0("V", row_number()))
mu_pred3a %>% as_data_frame() %>% 
  mutate(iter = 1:nrow(.)) %>% 
  pivot_longer(cols = -iter) %>% 
  left_join(times) %>% 
  group_by(iter) %>% 
  filter(value == max(value)) %>% 
  ungroup %>% 
  reframe(mean = mean(t4_seq),
          sd = sd(t4_seq))


tibble(mean = mu_mean3a,
       lower = mu_PI3a[1,],
       upper = mu_PI3a[2,],
       sd = sd(mu_mean3a),
       time = t4_seq) %>% 
  arrange(-mean)


mean_VO23 <- .339337 * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)
upper_VO23 <- (.339337 + 1.96 * (.3229105)) * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)
lower_VO23 <- (.339337 - 1.96 * (.3229105)) * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)


mean_VO23
upper_VO23
lower_VO23

mean_time3 <- 1.18139 * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)
upper_time3 <- (1.18139 + 1.96 * (.2941639)) * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)
lower_time3 <- (1.18139 - 1.96 * (.2941639)) * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)


mean_time3
upper_time3
lower_time3
```


#In this part I do the same thing again, this time withe the Thieli xp. I skipped doing the controls and the thieli 7p because i don't have as much data, and I think the controls wont be very interesting. Notably here, there is a pretty small standard deviation at the peak which we are 95% confident is somewhere between 3.17 and 4.17. This is lower than that for both of the Brygooi models though there is some overlap with the Brygooi xp. I'll also note that this model is not mass controlled, and the Brygooi are bigger, so it makes sense that they woudl have a higher peak. The peak here comes pretty quick though it does have a wide sd. Given the data, we can say that we are 95% confident that our peak lies somewhere between 213 and 1323 minutes. 

```{r}

mu_pred6 <- link(CO2_Time4, data = data.frame(T2 = t4_seq,
                                       STT = 8))
mu_mean6 <- apply(mu_pred6, 2, mean)
mu_PI6 <- apply(mu_pred6, 2, PI, prob = 0.95)

times = tibble(t4_seq = t4_seq) %>% 
  mutate(name = paste0("V", row_number()))
mu_pred6 %>% as_data_frame() %>% 
  mutate(iter = 1:nrow(.)) %>% 
  pivot_longer(cols = -iter) %>% 
  left_join(times) %>% 
  group_by(iter) %>% 
  filter(value == max(value)) %>% 
  ungroup %>% 
  reframe(mean = mean(t4_seq),
          sd = sd(t4_seq))

tibble(mean = mu_mean6,
       lower = mu_PI6[1,],
       upper = mu_PI6[2,],
       sd = sd(mu_mean6),
       time = t4_seq) %>% 
  arrange(-mean)


mean_VO26 <- -.2448264 * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)
upper_VO26 <- (-.2448264 + 1.96 * (.1281701)) * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)
lower_VO26 <- (-.2448264 - 1.96 * (.1281701)) * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)


mean_VO26
upper_VO26
lower_VO26

mean_time6 <- .05238649 * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)
upper_time6 <- (.0523864 + 1.96 * (.500217)) * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)
lower_time6 <- (.0523864 - 1.96 * (.500217)) * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)


mean_time6
upper_time6
lower_time6
```

#THis is where I finally graph the data. I found (through chatgpt) that the par() function is how to make facets in base R. I was able to change each label at the top of the panel to the actual trial type which was nice too. YOu can see in these graphs that the Brygooi 7p has a pretty strong curve, but I need more data for the ends. I also think I need to retake some of the fasted values for Thieli at the beginning because those are so high. There is also like no data for Thieli 7p. 


```{r}
# Set up a multi-panel layout: 2 rows, 3 columns (for 8 STT categories). I have no clue what the mar is but it's definitely necessary. 
par(mfrow = c(2, 4), mar = c(4, 4, 2, 1))

# X-axis sequence
t2_seq <- seq(-1.3, 2.6, length.out = 30)

stt_colors2 <- c("blue", "blue", "blue", "blue", "red", "red", "red", "red") 

# Define the real names for each STT category
stt_names <- c("Brygooi 7p", "Brygooi Fasted after 7p", "Brygooi Fasted after xp", "Brygooi xp", "Thieli 7p", "Thieli Fasted after 7p", "Thieli Fasted after xp", "Thieli xp")


# Loop through each STT category
for (stt in 1:8) {
  # Generate posterior predictions
  mu_pred <- link(CO2_Time4, data = data.frame(T2 = t2_seq, STT = stt))
  
  # Calculate mean and 95% CI
  mu_mean <- apply(mu_pred, 2, mean)
  mu_PI <- apply(mu_pred, 2, PI, prob = 0.95)
  
  # Filter observed data for this STT category
  observed_data <- sd_short %>% filter(STT == stt)
  
  # Plot observed data
  plot(observed_data$T2, observed_data$CO2,
       col = "black", pch = 16, cex = 0.7,
       xlab = "Time", ylab = "VCO2",
       main = stt_names[stt],
       ylim = range(c(observed_data$CO2, mu_PI)),
       xlim = range(t2_seq))
  
  # Add predicted mean line
  lines(t2_seq, mu_mean, col = stt_colors2[stt], lwd = 2)
  
  # Add shaded confidence intervals
  polygon(c(t2_seq, rev(t2_seq)),
          c(mu_PI[1, ], rev(mu_PI[2, ])),
          col = rgb(0, 0, 1, alpha = 0.2), border = NA)
}

# Reset layout to single panel
par(mfrow = c(1, 1))
```


#Here I wanted to look at just the 2 individuals (BthieM01 and BbrygM01) that make up the bulk of my data for each species and see how they compare. So, I just made this new dataset with just them. 

```{r}
sd_tidy2 <- sd_tidy %>%
  filter(Individual %in% c("BthieM01", "BbrygM01"))

sd_short2 = sd_tidy2 %>% dplyr::select(CO2, Species1, Species, T2, TrialType1, CPercent, TrialType, STT)
```


#This model is exactly the same as the other except it only includes data from the 2 individuals. 

```{r}
CO2_Time4a <-ulam(
alist(
  CO2 ~ dnorm(mu,sigma),
  #This includes an interaction between Time and Time squared since it is likely an upside down parabola. Each time you add the bracketed things it adds another line for that category
  mu <- a1[STT] + b1[STT] * T2 + b2[STT] * T2^2 ,
  c(a1, b1, b2)[STT] ~ multi_normal(c(a, b), Rho, sigma_id),
  a ~ dnorm(0, 1),
  b ~ dnorm(0, 1),
  sigma_id ~ dexp(1),
  Rho ~ lkj_corr(2),
 # b4 ~ dnorm(0, 1),
  sigma ~ dexp(1)
) ,data = sd_short2, chains = 4, cores = 4)
```



#Here you can compare the two models, and what I notice is that their predictions for the parameters are still quite close, probably because the new model is like 70% the same as the old model. Sort of expectedly, though , you can see that ess_bulk decreased when I took out more data. 

```{r}
precis(CO2_Time4a, depth = 3)
precis(CO2_Time4, depth = 3)
```


#Graphing this new model you can see that it is a little cleaner and the values (where I have them) are a bit tighter. I expected this to change some of the standard deviations that I was seeing in the previous model. 

```{r}
# Set up a multi-panel layout: 2 rows, 3 columns (for 6 STT categories)
par(mfrow = c(2, 4), mar = c(4, 4, 2, 1))

# X-axis sequence
t2_seq <- seq(-1.3, 2.6, length.out = 30)

stt_colors2 <- c("blue", "blue", "blue", "blue", "red", "red", "red", "red") 

# Define the real names for each STT category
stt_names <- c("Brygooi 7p", "Brygooi Fasted after 7p", "Brygooi Fasted after xp", "Brygooi xp", "Thieli 7p", "Thieli Fasted after 7p", "Thieli Fasted after xp", "Thieli xp")


# Loop through each STT category
for (stt in 1:8) {
  # Generate posterior predictions
  mu_pred <- link(CO2_Time4a, data = data.frame(T2 = t2_seq, STT = stt))
  
  # Calculate mean and 95% CI
  mu_mean <- apply(mu_pred, 2, mean)
  mu_PI <- apply(mu_pred, 2, PI, prob = 0.95)
  
  # Filter observed data for this STT category
  observed_data <- sd_short2 %>% filter(STT == stt)
  
  # Plot observed data
  plot(observed_data$T2, observed_data$CO2,
       col = "black", pch = 16, cex = 0.7,
       xlab = "Time", ylab = "VCO2",
       main = stt_names[stt],
       ylim = range(c(observed_data$CO2, mu_PI)),
       xlim = range(t2_seq))
  
  # Add predicted mean line
  lines(t2_seq, mu_mean, col = stt_colors2[stt], lwd = 2)
  
  # Add shaded confidence intervals
  polygon(c(t2_seq, rev(t2_seq)),
          c(mu_PI[1, ], rev(mu_PI[2, ])),
          col = rgb(0, 0, 1, alpha = 0.2), border = NA)
}

# Reset layout to single panel
par(mfrow = c(1, 1))
```


#For Mr. BrygooiM01 I notice that the values for the peak have shifted up a little bit from between 4 and 10 to now 5.3 and 11.6. This makes sense since we took out data from some smaller individuals (this is a big guy). The sd of time didnt change much, being about 30 mins from the peak on either side, though it is now a little earlier where it was between 986 and 1047 before it is now between 906 and 978 which is different, and we could say confidently a different prediction. There is probably then a sizeable effect of individual or at least body mass that this model is not accounting for. 

```{r}

mu_pred1a <- link(CO2_Time4a, data = data.frame(T2 = t4_seq,
                                       STT = 1))
mu_mean1a <- apply(mu_pred1a, 2, mean)
mu_PI1a <- apply(mu_pred1a, 2, PI, prob = 0.95)

times = tibble(t4_seq = t4_seq) %>% 
  mutate(name = paste0("V", row_number()))
mu_pred1a %>% as_data_frame() %>% 
  mutate(iter = 1:nrow(.)) %>% 
  pivot_longer(cols = -iter) %>% 
  left_join(times) %>% 
  group_by(iter) %>% 
  filter(value == max(value)) %>% 
  ungroup %>% 
  reframe(mean = mean(t4_seq),
          sd = sd(t4_seq))

tibble(mean = mu_mean1a,
       lower = mu_PI1a[1,],
       upper = mu_PI1a[2,],
       sd = sd(mu_mean1a),
       time = t4_seq) %>% 
  arrange(-mean)

mean_VO21a <- 2.171359897 * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)
upper_VO21a <- (2.171359897 + 1.96 * (0.7971939)) * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)
lower_VO21a <- (2.171359897 - 1.96 * (0.7971939)) * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)


mean_VO21a
upper_VO21a
lower_VO21a

mean_time1a <- 0.3592372 * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)
upper_time1a <- (0.3592372 + 1.96 * (0.03237605)) * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)
lower_time1a <- (0.3592372 - 1.96 * (0.03237605)) * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)


mean_time1a
upper_time1a
lower_time1a
```


# This part I take a look at the prediction for Mr. ThieliM01. The peak has shifted down from between 3.2 to 4.2 to now between 2.5 and 4.4. This is also a much larger CI which is probably due to the smaller amount of data. The time to peak has been noticeably changed where it is now a much longer time to the peak, between 1186 and 1518 minutes. This time range is much smaller ~ 300 minutes where the last time to peak prediction range was over 1000 minutes. This is probably because I took out the data from a large individual who I only had data for at the beginning which skewed the results. Interestingly, this value is now much longer than the Brygooi 7p where it had been a shorter time before. Also interesting, it now matches pretty closely to the values I have for Brygooi xp which peaks pretty late as well. I'm starting to think that the amount of food does have a big effect on the time to peak which is not something I've seen in published papers. 

```{r}

mu_pred6a <- link(CO2_Time4a, data = data.frame(T2 = t4_seq,
                                       STT = 8))
mu_mean6a <- apply(mu_pred6a, 2, mean)
mu_PI6a <- apply(mu_pred6a, 2, PI, prob = 0.95)

times = tibble(t4_seq = t4_seq) %>% 
  mutate(name = paste0("V", row_number()))
mu_pred6a %>% as_data_frame() %>% 
  mutate(iter = 1:nrow(.)) %>% 
  pivot_longer(cols = -iter) %>% 
  left_join(times) %>% 
  group_by(iter) %>% 
  filter(value == max(value)) %>% 
  ungroup %>% 
  reframe(mean = mean(t4_seq),
          sd = sd(t4_seq))

tibble(mean = mu_mean6a,
       lower = mu_PI6a[1,],
       upper = mu_PI6a[2,],
       sd = sd(mu_mean6a),
       time = t4_seq) %>% 
  arrange(-mean)

mean_VO26a <- -.3450226 * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)
upper_VO26a <- (-.3450226 + 1.96 * (0.2477631)) * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)
lower_VO26a <- (-.3450226 - 1.96 * (0.2477631)) * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)


mean_VO26a
upper_VO26a
lower_VO26a

mean_time6a <- 1.084313 * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)
upper_time6a <- (1.084313 + 1.96 * (.1494328)) * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)
lower_time6a <- (1.084313 - 1.96 * (.1494328)) * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)


mean_time6a
upper_time6a
lower_time6a
```



#Bonus model! In this dataset I added the variables logmass and CPercent to use for the model later.

```{r}
#Here I've made a new index variable which includes both Species and TrialType
sd_tidy = SDAmaster %>%
  mutate(SpeciesTrialType = paste(Species, TrialType, sep = "_"))


#This makes a dataset with all of the variables + New standardized Variables
sd_tidy3  = sd_tidy %>% 
  mutate(T2 = standardize(Time_Actual),
         STT = as.integer(as.factor(SpeciesTrialType)),
         CO2 = standardize(Average_VCO2),
         Species1 = as.integer(as.factor(Species)),
         TrialType1 = as.integer(as.factor(TrialType)),
         CPercent = standardize(Cricket_Percent),
         logmass = log(Body_Mass),
         logmass1 = standardize(logmass),
         TrialType,
         Species)

#This makes a much smaller dataset with only what you need which keeps the model from running the entire data set and lowers the amount of time it takes

sd_short3 = sd_tidy3 %>% dplyr::select(CO2, Species1, Species, T2, TrialType1, CPercent, TrialType, STT, logmass1)
```



#This model uses a couple extra parameters to predict our lines. I've added a standardized logmass variable and a CPercent which, while the mass of the crickets is somewhat controlled between conditions, there is some variation that will probably be accounted for in the xp prediction. 

```{r}

CO2_Time4b <-ulam(
alist(
  CO2 ~ dnorm(mu,sigma),
  #This includes an interaction between Time and Time squared since it is likely an upside down parabola. Each time you add the bracketed things it adds another line for that category
  mu <- a1[STT] + b1[STT] * T2 + b2[STT] * T2^2 + b3[STT] * logmass1 + b4[STT] * CPercent,
  c(a1, b1, b2, b3, b4)[STT] ~ multi_normal(c(a, b), Rho, sigma_id),
  a ~ dnorm(0, 1),
  b ~ dnorm(0, 1),
  sigma_id ~ dexp(1),
  Rho ~ lkj_corr(2),
  sigma ~ dexp(1)
) ,data = sd_short3, chains = 4, cores = 4)
```


# Using this model, you can see that there is a pretty high ess_bulk, and I notice that the same general trends are kept for parameters of the old model. Notably, the b2 is still negative and far from zero for STT- 1, 4, and 8 which is what I'd expect although 8 is now further from zero. We now see though that there is an effect of logmass, b3, on all but STT = 3, 5, and 7 which I have very little data for. This should help to account for the differences that I might find based on species size. Notably, I now don't see as much difference in the control values which means that we cannot say if one species has a greater mass-corrected metabolic rate than another which is interesting.  The parameter b4 which is cricket percent (the actual relative size of the meal rather than a binned version of 0, xp, or 7) has a large effect on both of the xp categories, STT= 4, and 8. This should help to make better predictions and put even though there is a little variation there. 

```{r}
precis(CO2_Time4b, depth = 3)
```


#Here are the graphs which look pretty similar to old ones. I was only able to run this if I took the mean values of logmass1 and CPercent which I don't really like, but I'll ask about this later. 

```{r}
# Set up a multi-panel layout: 2 rows, 3 columns (for 6 STT categories)
par(mfrow = c(2, 4), mar = c(4, 4, 2, 1))

# X-axis sequence
t2_seq <- seq(-1.3, 2.6, length.out = 30)

stt_colors2 <- c("blue", "blue", "blue", "blue", "red", "red", "red", "red") 

# Define the real names for each STT category
stt_names <- c("Brygooi 7p", "Brygooi Fasted after 7p", "Brygooi Fasted after xp", "Brygooi xp", "Thieli 7p", "Thieli Fasted after 7p", "Thieli Fasted after xp", "Thieli xp")


# Loop through each STT category
for (stt in 1:8) {
  # Generate posterior predictions
  mu_pred <- link(CO2_Time4b, data = data.frame(T2 = t2_seq, STT = stt,
    logmass1 = mean(sd_short3$logmass1[sd_short3$STT == stt], na.rm = TRUE),
    CPercent = mean(sd_short3$CPercent[sd_short3$STT == stt], na.rm = TRUE)))
  
  # Calculate mean and 95% CI
  mu_mean <- apply(mu_pred, 2, mean)
  mu_PI <- apply(mu_pred, 2, PI, prob = 0.95)
  
  # Filter observed data for this STT category
  observed_data <- sd_short3 %>% filter(STT == stt)
  
  # Plot observed data
  plot(observed_data$T2, observed_data$CO2,
       col = "black", pch = 16, cex = 0.7,
       xlab = "Time", ylab = "VCO2",
       main = stt_names[stt],
       ylim = range(c(observed_data$CO2, mu_PI)),
       xlim = range(t2_seq))
  
  # Add predicted mean line
  lines(t2_seq, mu_mean, col = stt_colors2[stt], lwd = 2)
  
  # Add shaded confidence intervals
  polygon(c(t2_seq, rev(t2_seq)),
          c(mu_PI[1, ], rev(mu_PI[2, ])),
          col = rgb(0, 0, 1, alpha = 0.2), border = NA)
}

# Reset layout to single panel
par(mfrow = c(1, 1))
```


# Interestingly, here we find that there is a big difference between the prediction for the peak VCO2 when mass corrected for the Brygooi. Where it used to predict somewhere between 5 and 11, we now see values which have a lower standard deviation between 2.2 and 4.2. Expectedly, the time to peak hasnt changed much and is still right around the 1000 mark. 

```{r}

mu_pred1b <- link(CO2_Time4b, data = data.frame(T2 = t4_seq,
                                       STT = 1,
    logmass1 = mean(sd_short3$logmass1[sd_short3$STT == stt], na.rm = TRUE),
    CPercent = mean(sd_short3$CPercent[sd_short3$STT == stt], na.rm = TRUE)))
mu_mean1b <- apply(mu_pred1b, 2, mean)
mu_PI1b <- apply(mu_pred1b, 2, PI, prob = 0.95)

times = tibble(t4_seq = t4_seq) %>% 
  mutate(name = paste0("V", row_number()))
mu_pred1b %>% as_data_frame() %>% 
  mutate(iter = 1:nrow(.)) %>% 
  pivot_longer(cols = -iter) %>% 
  left_join(times) %>% 
  group_by(iter) %>% 
  filter(value == max(value)) %>% 
  ungroup %>% 
  reframe(mean = mean(t4_seq),
          sd = sd(t4_seq))

tibble(mean = mu_mean1b,
       lower = mu_PI1b[1,],
       upper = mu_PI1b[2,],
       sd = sd(mu_mean1b),
       time = t4_seq) %>% 
  arrange(-mean)

mean_VO21b <- -.4636748 * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)
upper_VO21b <- (-.4636748 + 1.96 * (-0.2552429)) * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)
lower_VO21b <- (-.4636748 - 1.96 * (-0.2552429)) * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)


mean_VO21b
upper_VO21b
lower_VO21b

mean_time1b <- 0.4633212 * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)
upper_time1b <- (0.4633212 + 1.96 * (0.02326078)) * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)
lower_time1b <- (0.4633212 - 1.96 * (0.02326078)) * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)


mean_time1b
upper_time1b
lower_time1b
```

# With our mass-corrected model we now see that Thieli has a little bit higher predicted peak, while there is a good amount of overlap, we are 95% sure that Thieli is between 3.15 and 4.67. This peak also now has no overlap with the peak of the Brygooi where the range of the CI is between 1102 and 1452 minutes. Given this difference, we can say that we are confident that it takes the Thieli longer to reach the peak and has a higher one at that.


```{r}

mu_pred8b <- link(CO2_Time4b, data = data.frame(T2 = t4_seq,
                                       STT = 8,
    logmass1 = mean(sd_short3$logmass1[sd_short3$STT == stt], na.rm = TRUE),
    CPercent = mean(sd_short3$CPercent[sd_short3$STT == stt], na.rm = TRUE)))

mu_mean8b <- apply(mu_pred8b, 2, mean)
mu_PI8b <- apply(mu_pred8b, 2, PI, prob = 0.95)

times = tibble(t4_seq = t4_seq) %>% 
  mutate(name = paste0("V", row_number()))
mu_pred8b %>% as_data_frame() %>% 
  mutate(iter = 1:nrow(.)) %>% 
  pivot_longer(cols = -iter) %>% 
  left_join(times) %>% 
  group_by(iter) %>% 
  filter(value == max(value)) %>% 
  ungroup %>% 
  reframe(mean = mean(t4_seq),
          sd = sd(t4_seq))

tibble(mean = mu_mean8b,
       lower = mu_PI8b[1,],
       upper = mu_PI8b[2,],
       sd = sd(mu_mean8b),
       time = t4_seq) %>% 
  arrange(-mean)

mean_VO28b <- -0.1248532 * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)
upper_VO28b <- (-0.1248532 + 1.96 * (0.1941774)) * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)
lower_VO28b <- (-0.1248532 - 1.96 * (0.1941774)) * sd(SDAmaster$Average_VCO2) + mean(SDAmaster$Average_VCO2)


mean_VO28b
upper_VO28b
lower_VO28b

mean_time8b <- 0.9514809 * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)
upper_time8b <- (0.9514809 + 1.96 * (0.1581041)) * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)
lower_time8b <- (0.9514809 - 1.96 * (0.1581041)) * sd(SDAmaster$Time_Actual) + mean(SDAmaster$Time_Actual)


mean_time8b
upper_time8b
lower_time8b
```


# I think moving forward with this data I am going to do a percent change where I go through and calculate Standard Metabolic Rate (SMR) from control data and then use that to find a percent change in MR over time. I'll then look to see where that line crosses 0 to predict the total duration of the increase in MR. Using that, I'll look to calculate the area under that curve to find total cost of digestion for each individual/species. All of this is to say that I'll probably be in your office asking for help!


# Lastly, I wanted to say thank you for a great semester, and that I feel like I learned a lot! I'm not sure if I'll be in Bayes next semester, but I'll definitely be using it for my projects going foward. Thanks again!!

# Gene
