---
title: "SDA_brms"
output: html_document
date: "2025-01-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r}
#install.packages("brms")
#install.packages("janitor")
#install.packages("tidyverse")
#install.packages("tidybayes")
#install.packages("rethinking")

library(brms)
library(janitor)
library(tidyverse)
library(tidybayes)
library(rethinking)
```

#brms

brmModel <- brm(Petal.Length ~ Sepal.Length,
                , data = iris,
                family = Gamma(ling = "log"),
                prior = c(prior(normal(0,5), class = b),
                          prior(normal(-5, 2), class = Intercept),
                          prior(exponential(1), class = sigma)))

brmModel2 <- brm(Petal.Length ~ 1 + Sepal.Length + (1 + Sepal.Length|Species),
                , data = iris,
                family = Gamma(link = "log"))

saveRDS(brmModel2, file = "Bayesian/brmModel2.rds")

brmModel2 - readRDS(brmModel2, file = "Bayesian/brmModel2.rds")

summary(brmModel)
iris_clean = iris %>% clean_names()


mod_brm_update = update(brmModel, newdata = iris_clean,
                        formula = petal_length ~ sepal_length,
                         prior = c(prior(normal(0,5), class = b),
                          prior(normal(-5, 2), class = Intercept),
                          prior(exponential(1), class = sigma)))

plot_data <- plot(conditional_effects(brmModel2), points = T)

plot_data$Sepal.Length:Species +
  labs(y = "Petal Length (mm)",
       x = "Sepal Length (mm)")


```{r}
#Here I've made a new index variable which includes both Species and TrialType
sd_tidy = SDAmaster %>%
  mutate(SpeciesTrialType = paste(Species, TrialType, sep = "_"))

sd_tidy_heros <- sd_tidy %>%
  filter(Individual == c("BthieM01", "BbrygM02"))


#This makes a dataset with all of the variables + New standardized Variables
sd_tidy_heros3  = sd_tidy_heros %>% 
  mutate(T2 = standardize(Time_Actual),
         STT = as.integer(as.factor(SpeciesTrialType)),
         CO2 = standardize(Average_VCO2),
         Species1 = as.integer(as.factor(Species)),
         TrialType1 = as.integer(as.factor(TrialType)),
         CPercent = standardize(Cricket_Percent),
         logmass = log(Body_Mass),
         logmass1 = standardize(logmass),
         TrialType,
         Species)

#This makes a much smaller dataset with only what you need which keeps the model from running the entire data set and lowers the amount of time it takes

sd_short_heros3 = sd_tidy_heros3 %>% dplyr::select(CO2, Species1, Species, T2, TrialType1, CPercent, TrialType, STT, logmass1)
```


```{r}


brmModel2 <- brm(CO2 ~ 1 + T2 + (1 + T2|STT),
                , data = sd_short_heros3,
                family = gaussian(link = "log"))

```


```{r}

```


```{r}
plot_data <- plot(conditional_effects(brmModel2), points = T)

plot_data$T2:STT +
  labs(y = "Petal Length (mm)",
       x = "Sepal Length (mm)")
```


```{r}
CO2_Time4b_brms <- brm(
  formula = CO2 ~ 1 + T2 + I(T2^2) + logmass1 + CPercent +  # Fixed effects
    (1 + T2 + I(T2^2) + logmass1 + CPercent || STT),        # Group-level effects with separate variances
  data = sd_short_heros3,
  family = gaussian()
)
```


```{r}
plot(CO2_Time4b_brms)
```

```{r}
summary(CO2_Time4b_brms)
```

```{r}
fixef(CO2_Time4b_brms)
```

```{r}
# Random effects for each STT category
ranef(CO2_Time4b_brms)

# Combined effects (fixed + random)
coef(CO2_Time4b_brms)

# Marginal effects plots
conditional_effects(CO2_Time4b_brms, effects = "STT")
```

```{r}
bayes_R2(CO2_Time4b_brms)
```








,                                      # Normal distribution
  prior = c(
    # Priors for fixed effects
    prior(normal(0, 1), class = "Intercept"),                # Prior for intercept
    prior(normal(0, 1), class = "b"),                        # Prior for slopes (T2, T2^2, etc.)
    
    # Priors for group-level effects
    prior(exponential(1), class = "sd"),                     # Prior for standard deviations
    prior(lkj(2), class = "cor")                             # LKJ prior for correlation matrix
  ),
  chains = 4, cores = 4, iter = 2000, seed = 123
)

```{r}
saveRDS(CO2_Time4b_brms, file = "C:/Users/19glo/OneDrive - The University of South Dakota/SDA/SDA/CO2_Time4b_brms.rds")

```

```{r}
# Set up a multi-panel layout: 2 rows, 3 columns (for 6 STT categories)
par(mfrow = c(2, 4), mar = c(4, 4, 2, 1))

# X-axis sequence
t2_seq <- seq(-1.3, 2.6, length.out = 30)

stt_colors2 <- c("blue", "blue", "blue", "blue", "red", "red", "red", "red") 

# Define the real names for each STT category
stt_names <- c("Brygooi 7p", "Brygooi Fasted after 7p", "Brygooi Fasted after xp", "Brygooi xp", "Thieli 7p", "Thieli Fasted after 7p", "Thieli Fasted after xp", "Thieli xp")


# Loop through each STT category
for (stt in 1:8) {
  # Generate posterior predictions
  mu_pred <- link(CO2_Time4b_brms, data = data.frame(T2 = t2_seq, STT = stt,
    logmass1 = mean(sd_short3$logmass1[sd_short3$STT == stt], na.rm = TRUE),
    CPercent = mean(sd_short3$CPercent[sd_short3$STT == stt], na.rm = TRUE)))
  
  # Calculate mean and 95% CI
  mu_mean <- apply(mu_pred, 2, mean)
  mu_PI <- apply(mu_pred, 2, PI, prob = 0.95)
  
  # Filter observed data for this STT category
  observed_data <- sd_short3 %>% filter(STT == stt)
  
  # Plot observed data
  plot(observed_data$T2, observed_data$CO2,
       col = "black", pch = 16, cex = 0.7,
       xlab = "Time", ylab = "VCO2",
       main = stt_names[stt],
       ylim = range(c(observed_data$CO2, mu_PI)),
       xlim = range(t2_seq))
  
  # Add predicted mean line
  lines(t2_seq, mu_mean, col = stt_colors2[stt], lwd = 2)
  
  # Add shaded confidence intervals
  polygon(c(t2_seq, rev(t2_seq)),
          c(mu_PI[1, ], rev(mu_PI[2, ])),
          col = rgb(0, 0, 1, alpha = 0.2), border = NA)
}

# Reset layout to single panel
par(mfrow = c(1, 1))
```

